---
title: "R2 Matrix"
author: "LV"
date: "2020-08-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Load Libraries
Run in R:
```{r}
suppressPackageStartupMessages(library(RSQLite))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(data.table))
```


```{bash, eval=FALSE}
cd $DATA
mkdir PEC_TWAS_weights
cd PEC_TWAS_weights
wget "http://resource.psychencode.org/Datasets/Derived/PEC_TWAS_weights.tar.gz"
tar xvf PEC_TWAS_weights.tar.gz
```


# Definitions
```{r}
PRE="/Users/lvairus"
DATA=glue::glue("{PRE}/data-Github/psychencode")
OUTPUT=DATA
```


This script is to copy and paste the rsq data from the PEC_TWAS_weights.tar.gz files into a matrix
```{r read r2 files into mat, eval=FALSE}
filelist <- list.files(glue::glue("{DATA}/PEC_TWAS_weights/"))
filelist <- filelist[substr(filelist,1,4)=="ENSG"]

file=filelist[1]

load(glue::glue("{DATA}/PEC_TWAS_weights/{file}"))
ngenes <- length(filelist)
R2_mat <- matrix(NA, ngenes, 5)
genelist <- substr(filelist, 1, 15)
rownames(R2_mat) <- genelist
colnames(R2_mat) <- colnames(cv.performance)


for (file in filelist){
  load(glue::glue("{DATA}/PEC_TWAS_weights/{file}"))
  genename <- substr(file, 1, 15)
  R2_mat[genename, ] <- cv.performance["rsq", ]
}

saveRDS(R2_mat, "~/data-Github/psychencode/R2_mat.RDS")

```



```{r Comparison of Models}

if( !exists("R2_mat") ) { 
  R2_mat = readRDS(glue::glue("{DATA}/R2_mat.RDS"))
} 

plot(R2_mat[,"enet"],R2_mat[,"top1"],main="CV R2 comparison top SNP vs enet");abline(0,1,col='gray')
plot(R2_mat[,"enet"],R2_mat[,"blup"],main="CV R2 comparison blup vs enet");abline(0,1,col='gray')
plot(R2_mat[,"enet"],R2_mat[,"bslmm"],main="CV R2 comparison bslmm vs enet");abline(0,1,col='gray')
plot(R2_mat[,"enet"],R2_mat[,"lasso"],main="CV R2 comparison lasso vs enet");abline(0,1,col='gray')

pairs(R2_mat)

```

## Conclusion: these figure shows that Elastic Net has overall better prediction performance (points are above the one to one line)
You can see on the graphs that, overall, the prediction performance of the Elastic Net is larger compared to the other methods.


