---
title: "psychencode_hg38_validation"
author: ""
date: "2020-12-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


# Liftover
## Definitions

```{bash, eval=FALSE}
conda activate imlabtools
PRE=/Users/sabrinami/Github/analysis-sabrina/prediction_model_liftover
METAXCAN=/Users/sabrinami/Github/MetaXcan/software
DATA=/Users/sabrinami/Desktop/psychencode_test_data
MODEL=$PRE/models
LIFTOVER=$PRE/liftover
RESULTS=$PRE/results
cd $PRE
```

## Model DB
Run the script with the PrediXcan format PsychENCODE model, and a liftover file which can be downloaded: `wget https://hgdownload.soe.ucsc.edu/goldenPath/hg19/liftOver/hg19ToHg38.over.chain.gz`
```{bash, eval=FALSE}
python prediction_model_liftover.py \
--input_model_db models/psychencode.db \
--liftover liftover/hg19ToHg38.over.chain.gz \
--output_model_db models/psychencode_hg38.db
```

Some of the variants in the hg19 model failed to lift over to hg38, they are recorded in `dropped_snps.csv`

## Generate Covariances
A more complete tutorial is here: https://github.com/hakyimlab/analysis-sabrina/blob/master/covariance_1000G_ref/covariances_hg38.Rmd.
I submitted this job in CRI: `/gpfs/data/im-lab/nas40t2/sabrina/scripts/calculate_covariance_1000G_hg38/psychencode_cov_1000G_hg38.sh`

# Validation
## Definitions

Similarly, in R, load the libraries, then set the same definitions.
```{r libraries}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(qqman))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(RSQLite))
suppressPackageStartupMessages(library(UpSetR))
```


```{r definitions}
PRE="/Users/sabrinami/Github/analysis-sabrina/prediction_model_liftover"
DATA="/Users/sabrinami/Desktop/psychencode_test_data"
RESULTS=glue::glue("{PRE}/results")
MODEL=glue::glue("{PRE}/models")
CODE="/Users/sabrinami/Github/psychencode/code"
source(glue::glue("{CODE}/load_data_functions.R"))
source(glue::glue("{CODE}/plotting_utils_functions.R"))

gencode_df = load_gencode_df()
```

## Run S-PrediXcan

Run S-PrediXcan on the original model in hg19.
```{bash, eval=FALSE}
python3 $METAXCAN/SPrediXcan.py --gwas_file $DATA/clozuk_pgc2.meta.sumstats.out.txt \
--model_db_path $MODEL/psychencode.db \
--covariance $MODEL/psychencode_varID.txt.gz \
--snp_column varID \
--or_column OR \
--pvalue_column P \
--non_effect_allele_column A2 \
--effect_allele_column A1 \
--keep_non_rsid --additional_output --model_db_snp_key varID \
--throw \
--output_file $RESULTS/clozuk_pgc2_psychencode.csv
```


Run S-PrediXcan with the lifted over model.
```{bash, eval=FALSE}
python $METAXCAN/SPrediXcan.py \
--gwas_file  $DATA/imputed_clozuk_pgc2.txt.gz \
--snp_column panel_variant_id --effect_allele_column effect_allele --non_effect_allele_column non_effect_allele --zscore_column zscore \
--model_db_path $MODEL/psychencode_hg38.db \
--covariance $MODEL/psychencode_hg38.txt.gz \
--keep_non_rsid --additional_output --model_db_snp_key varID \
--throw \
--output_file $RESULTS/clozuk_pgc2_psychencode_hg38.csv
```

And the mashr model.
```{bash, eval=FALSE}
python $METAXCAN/SPrediXcan.py \
--gwas_file  $DATA/imputed_clozuk_pgc2.txt.gz \
--snp_column panel_variant_id --effect_allele_column effect_allele --non_effect_allele_column non_effect_allele --zscore_column zscore \
--model_db_path $MODEL/mashr_Brain_Cortex.db \
--covariance $MODEL/mashr_Brain_Cortex.txt.gz \
--keep_non_rsid --additional_output --model_db_snp_key varID \
--throw \
--output_file $RESULTS/clozuk_pgc2_mashr_Brain_Cortex.csv
```

## Compare Association Results

First, load all assocation results, and check for significant genes.
```{r load hg19 association}
spredixcan_association_psychencode_hg19 = load_spredixcan_association(glue::glue("{RESULTS}/clozuk_pgc2_psychencode.csv"), gencode_df)
dim(spredixcan_association_psychencode_hg19)
significant_genes_psychencode_hg19 <- spredixcan_association_psychencode_hg19 %>% filter(pvalue < 0.05/nrow(spredixcan_association_psychencode_hg19)) %>% arrange(pvalue)
```

```{r load hg38 association}
spredixcan_association_psychencode_hg38 = load_spredixcan_association(glue::glue("{RESULTS}/clozuk_pgc2_psychencode_hg38.csv"), gencode_df)
dim(spredixcan_association_psychencode_hg38)
significant_genes_psychencode_hg38 <- spredixcan_association_psychencode_hg38 %>% filter(pvalue < 0.05/nrow(spredixcan_association_psychencode_hg38)) %>% arrange(pvalue)
```

```{r load mashr association}
spredixcan_association_Brain_Cortex = load_spredixcan_association(glue::glue("{RESULTS}/clozuk_pgc2_mashr_Brain_Cortex.csv"), gencode_df)
dim(spredixcan_association_Brain_Cortex)
significant_genes_Brain_Cortex <- spredixcan_association_Brain_Cortex %>% filter(pvalue < 0.05/nrow(spredixcan_association_Brain_Cortex)) %>% arrange(pvalue)
```

As a sanity check, check PsychENCODE models are consistent between builds.

```{r compare psychencode zscores}
psychencode_zscores = inner_join(spredixcan_association_psychencode_hg19, spredixcan_association_psychencode_hg38, by=c("gene"))
dim(psychencode_zscores)
psychencode_zscores %>% ggplot(aes(zscore.x, zscore.y)) + geom_point() + ggtitle("PsychENCODE S-PrediXcan z-score") + xlab("hg19") + ylab("hg38") + geom_abline(intercept = 0, slope = 1)
```

Then compare PsychENCODE and mashr Brain Cortex z-scores.

```{r compare psychencode mashr zscores}
psychencode_Brain_Cortex_zscores = inner_join(spredixcan_association_Brain_Cortex, spredixcan_association_psychencode_hg38, by=c("gene"))
dim(psychencode_Brain_Cortex_zscores)
psychencode_zscores %>% ggplot(aes(zscore.x, zscore.y)) + geom_point() + ggtitle("S-PrediXcan z-score") + xlab("mashr Brain Cortex") + ylab("PsychENCODE") + geom_abline(intercept = 0, slope = 1)
```


